---
title: 'Exercise: Linear algebra'
author: "Jakob Liebmann"
date: "15.05.2020"
output: 
  html_document:
    toc: true
    toc_float: true
    df_: paged
    code_folding: show  
    theme: united
    highlight: tango
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1

Consider the following three vectors 
- a = (−0.5276,−0.8453,−0.0839) 
- b = (−0.6050, 0.4433,−0.6614) 
- c = (−0.5963, 0.2981, 0.7454)
```{r}
a <- c(−0.5276,−0.8453,−0.0839)
b <- c(−0.6050, 0.4433,−0.6614)
c <- c(−0.5963, 0.2981, 0.7454)
```


Calculate the Euclidean length of a, b, and c. 
```{r}
# writing a function doing the job
length_vector <- function(vector){
  square <- 0
  for (i in vector){
    square <- square + (i*i)
  }
  length_vec = sqrt(square)
  return(length_vec)
}

length_vector(a)
length_vector(b)
length_vector(c)
```


Then calculate the dot products a · b, a · c, and b · c.
```{r}
a %*% b
a %*% c
b %*% c
```


What is the name for such vectors?

Because they are unit vectors (length = 1) and orthogonal vectors (dot product = 0) at the same time, these vectors are called orthonormal.

# Exercise 2

Consider the following two vectors: 
u = (0.5, 0.4, 0.4, 0.5, 0.1, 0.4, 0.1) and 
v = (−1,−2, 1,−2, 3, 1,−5).
```{r}
u <- c(0.5, 0.4, 0.4, 0.5, 0.1, 0.4, 0.1)
v <- c(−1,−2, 1,−2, 3, 1,−5)
```


• Calculate the Euclidean length of vectors u using the dot product notation. How do you call a vector of this length?
```{r}
u %*% v
```


• Calculate the Euclidean length of vector v. Then create v_normalized by dividing each element of v by the length of v. Then check the length of v_normalized.
```{r}
length_vector(v)
v_normalized <- v/length_vector(v)
length_vector(v_normalized)
```


• Calculate the outer product of u and v, an their dot product. What is the geometric interpration of the dot product?
```{r}
# outer product
u %*% t(v)

# inner (dot) product
u %*% v
```
The inner/dot product describes how often vector a fits in the projection of vector b on vector a.


• Calculate the angle between u and v using the fact that u · v = ||u|| ||v|| cos(theta). 
(Hint: the base R function acos does not return “degrees”, but “radians”. You can convert radians to degrees by multiplying with 180/pi. As an alternative you can use the function aspace::acos_d).
```{r}
theta = acos((u %*% v) / (length_vector(u) * length_vector(v)))
(theta <- theta * 180 / pi)
```


# Exercise 3
A shop sells apples, cherries, and bananas for a price of 1€, 3€, and 2€ respectively per kilogram. The following table (see pdf-file) shows the amounts (in kg) of apples, cherries, and bananas sold in the course of a week.

• Turn the table into a matrix and use matrix-vector multiplication to calculate the value of total sales per day.
```{r}
(shop <-matrix(c(5,4,3,6,10,1,2,2,4,5,4,4,3,6,3), nrow = 3, ncol = 5, byrow = TRUE))
(prices <- c(1,3,2))
prices %*% shop
```

• What characterises linear dependence of vectors? Do you see linear dependent columns?

Linear dependence of column-vectors descripes redundant information within one column. I don't see this case.

• Is the following statement true: This matrix can have at most three linearly independent columns.

Thats true, the number of dimensions is limitting the maximum amount of independent vectors.
